---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---

## Loading libraries
```{r, results="hide"}
## Load libraries
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggplot2)
```

## Loading and preprocessing the data
```{r}
unzip ("./activity.zip")
file <- read.csv("./activity.csv")
ds1 <- file %>% mutate(date = as.Date(date))
head(ds1)
```


## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day.
```{r hist1}
ds2 <- ds1 %>% group_by(date) %>% summarize(total = sum(steps, na.rm = TRUE))
head(ds2)

qplot(total, data = ds2, main = "Histogram of steps taken each day")
```

Mean and median total number of steps taken per day.
```{r}
ds2 %>% summarize(avg = mean(total, na.rm = TRUE), med = median(total, na.rm = TRUE))
```


## What is the average daily activity pattern?
Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r timeseries1}
ds4 <- ds1 %>% group_by(interval) %>% summarize(avg = mean(steps, na.rm = TRUE))
head(ds4)

qplot(interval, avg, data = ds4, geom = "line", ylab = "Steps",
      main = "Average number of steps taken")
```

5-minute interval, on average across all the days in the dataset, with the maximum number of steps.
```{r}
ds4 %>% filter(avg == max(ds4$avg))
```


## Imputing missing values
Total number of missing values in the dataset.
```{r}
## Missing values in the dataset
sum(is.na(ds1$steps))
```

Select median for the 5-min interval to fill in values.
```{r}
## Median for each 5-min interval
ds5 <- ds1 %>% group_by(interval) %>% summarize(med = median(steps, na.rm = TRUE))
head(ds5)
```

Time series plot of median 5-min interval.
```{r timeseries2}
## Time series plot
qplot(interval, med, data= ds5, geom = "line", ylab = "Steps",
      main = "Median 5-minute interval")
```

Filling in missing values with the median for the 5-min interval.
```{r}
## Missing values filling in
ds6 <- ds1 %>% group_by(interval) %>% mutate(med = median(steps, na.rm = TRUE)) %>%
    mutate(steps_med = ifelse(is.na(steps), med, steps))
ds7 <- ds6 %>% select(date, interval, steps_med)
head(ds7)
```

Histogram of the total number of steps taken each day.
```{r hist2}
## Number of steps taken each day
ds8 <- ds7 %>% group_by(date) %>% summarize(total = sum(steps_med))
head(ds8)

qplot(total, data = ds8, main = "Histogram of steps taken each day")
```

Mean and median total number of steps taken per day.
```{r}
ds8 %>% summarize(avg = mean(total), med = median(total))
```
Mean is higher. Median keeps the same.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## Set locale
Sys.setlocale("LC_TIME", locale = "en_US.UTF-8")

## Weekdays
ds10 <- ds7 %>% mutate(weekday = weekdays(date))

## Weekdays as weekend, weekday
ds11 <- ds10 %>% mutate(wd = as.factor(
    ifelse(weekday %in% c("Saturday", "Sunday"), "weekend", "weekday")))
head(ds11)
```

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days.
```{r timeseries3}
## Averaged across all weekday days or weekend
ds12 <- ds11 %>% group_by(interval, wd) %>% summarize(avg = mean(steps_med))
head(ds12)

qplot(interval, avg, data = ds12, facets = wd ~ ., geom = "line", ylab = "Steps",
      main = "Average number of steps taken")
```
